LCA in R using tidySEM

### 2022-09-14 example_analysis/Daniel_LCA.Rmd
Update 14 Septemeber
In the manuscript.Rmd:

Added Shapiro-Wilks test when discussing Spurk.
Removed some obsolete comments (we still have backup just in case).
Wrote about alternative model parametrizations in LCA, referring back to GRoLTS.
Under Model Fit Indices, I wrote about -2*LL, mentioned we are working on an implementation of the lazy bootstrap.
Finished the Classification Diagnostics and Interpreting Class Solution sections.

In Daniel_LCA:

Finished annotating the code. We can include this section in the manuscript in the next commit.

---
title: "LCA in R using tidySEM"
output: html_notebook
---

# LOADING THE DATASET

```{r}
library(tidySEM) 
df <- data_mix_ordinal #ask Caspar to rename the dataset in documentation
df[1:4] <- lapply(df, ordered)
```

# Exploring the data

```{r}
tidySEM::descriptives(df)
```

# Conduct latent class analysis

```{r}
# fitting the model
set.seed(123)
res <- mx_lca(data = df, classes = 1:4)
```
# Class enumeration

```{r}
# printing model fit table
fit_table <- table_fit(res)
fit_table[ , c("Name", "LL", "AIC", "BIC", "Entropy", "prob_min", "prob_max", "n_min", "n_max")]
```

# Interpreting class solution

```{r}
library(ggplot2)
plot_prob(res[[3]]) + scale_fill_manual(name = "bla")
table_prob(res[[3]])
```

@@

```{r}
probs <- class_prob(res[[3]])
probs$mostlikely.class
probs$individual
```

# APPENDIX

USING DIFFERENT PACKAGES

This is an example of a Latent Class Analysis (LCA).

In Part 1, I use the Lazarsfeld dataset and OpenMX.
In Part 2, I use the Drinking dataset and OpenMx.
In Part 3, I redo the analysis from Part 2 using tidySEM.

```{r}
library(tidySEM)
library(OpenMx)
```

Part 1: LCA using OpenMx and Lazarsfeld Dataset

Analysis based on: https://rdrr.io/cran/OpenMx/src/inst/models/passing/LCAlazarsfeld.R

```{r}
data("lazarsfeld", package ="OpenMx", verbose= TRUE)
data <- lazarsfeld

freq <- data[,5]
data[,1] <- as.ordered(data[,1])
data[,2] <- as.ordered(data[,2])
data[,3] <- as.ordered(data[,3])
data[,4] <- as.ordered(data[,4])
vars <- data[,1:4]

nclass <- 2
nvar <- 4
nthresh <- 1

nameList <- names(vars)

class1 <- mxModel("Class1", 
            mxMatrix("Iden", name = "R", nrow = nvar, ncol = nvar, free=FALSE),
            mxMatrix("Full", name = "M", nrow = 1, ncol = nvar, free=FALSE),
            mxMatrix("Full", name = "ThresholdsClass1", nrow = 1, ncol = nvar, 
            	dimnames = list("Threshold",nameList), free=TRUE),
		  mxExpectationNormal(covariance="R", means="M", dimnames=nameList, 
				      thresholds="ThresholdsClass1"),
		  mxFitFunctionML(vector=TRUE))

class2 <- mxModel("Class2", 
            mxMatrix("Iden", name = "R", nrow = nvar, ncol = nvar, free=FALSE),
            mxMatrix("Full", name = "M", nrow = 1, ncol = nvar, free=FALSE),
            mxMatrix("Full", name = "ThresholdsClass2", nrow = 1, ncol = nvar, 
            	dimnames = list("Threshold",nameList), free=TRUE),
		  mxExpectationNormal(covariance="R", means="M", dimnames=nameList,
				      thresholds="ThresholdsClass2"),
		  mxFitFunctionML(vector=TRUE))

# Define the model

lcamodel <- mxModel("lcamodel", class1, class2, mxData(vars, type="raw"),

# Create class membership probabilities, constrain them to be in the range 0-1, and make their sum equal 1.0
            mxMatrix("Full", name = "ClassMembershipProbabilities", nrow = nclass, ncol = 1, free=TRUE, 
			        labels = c(paste("pclass", 1:nclass, sep=""))),
            mxBounds(c(paste("pclass", 1:nclass, sep="")),0,1),
		    mxMatrix("Iden", nrow = 1, name = "constraintLHS"),
		    mxAlgebra(sum(ClassMembershipProbabilities), name = "constraintRHS"),
            mxConstraint(constraintLHS == constraintRHS),

# Define the objective function
            mxAlgebra(-2*sum(freq * log(pclass1%x%Class1.objective + pclass2%x%Class2.objective)), 
            	name="lca"),
            mxFitFunctionAlgebra("lca")
)

# Run the job
model <- mxRun(lcamodel, suppressWarnings=TRUE)
summary(model)
model$matrices

# Check results against those hard-coded from old Mx:
negativeThresholds <- as.vector(mxEval(Class1.ThresholdsClass1, model))
positiveThresholds <- as.vector(mxEval(Class2.ThresholdsClass2, model))
if (negativeThresholds[[1]] > 0) {
	temp <- negativeThresholds
	negativeThresholds <- positiveThresholds
	positiveThresholds <- temp
}

omxCheckCloseEnough(negativeThresholds, c(-1.3247, -0.2909, -0.1466, -0.0044),.01)
omxCheckCloseEnough(positiveThresholds, c(0.1805, 0.9071, 1.3169, 1.5869),.01)
omxCheckCloseEnough(sort(as.vector(mxEval(ClassMembershipProbabilities, model))), c(0.4440, 0.5560),.01)
omxCheckCloseEnough(model$output$Minus2LogLikelihood, 4696.444, 0.01)
```

Part 2: LCA Using OpenMx and Drinking Dataset,

Dataset from: https://stats.oarc.ucla.edu/sas/dae/latent-class-analysis/

```{r}
data <- read.csv('C:/Users/danie/OneDrive/Desktop/LCA/drinking_data.csv', header = T, sep = ',')

#check if two rows have the same response pattern
library(DAKS)
pattern(data)

dat1 <- c(1,0,0,0,0,0,0,0,0, 126)
dat2 <- c(0,0,0,0,0,0,0,0,0, 102)
dat3 <- c(1,0,0,0,0,0,0,0,1, 47)
dat4 <- c(1,1,0,0,0,0,0,0,0, 47)
dat5 <- c(1,0,0,0,0,1,0,0,0, 43)
dat6 <- c(126, 102, 47, 47, 43)

dat <- data.frame(dat1,dat2,dat3,dat4,dat5)
dat <- t(dat)
dat <- as.data.frame(dat)
names(dat) <- c("Q1", "Q2", "Q3", "Q4", "Q5", "Q6", "Q7", "Q8", "Q9", "freq")
rm("data", "dat1", "dat2", "dat3", "dat4", "dat5", "dat6")

dat[,1] <- as.ordered(dat[,1])
dat[,2] <- as.ordered(dat[,2])
dat[,3] <- as.ordered(dat[,3])
dat[,4] <- as.ordered(dat[,4])
dat[,5] <- as.ordered(dat[,5])
dat[,6] <- as.ordered(dat[,6])
dat[,7] <- as.ordered(dat[,7])
dat[,8] <- as.ordered(dat[,8])
dat[,9] <- as.ordered(dat[,9])

freq <- dat[,10]
vars <- dat[,1:9]

nclass <- 2
nvar <- 9
nthresh <- 1
nameList <- names(vars)

class1 <- mxModel("Class1", 
                  mxMatrix("Iden", name = "R", nrow = nvar, ncol = nvar, free=FALSE),
                  mxMatrix("Full", name = "M", nrow = 1, ncol = nvar, free=FALSE),
                  mxMatrix("Full", name = "ThresholdsClass1", nrow = 1, ncol = nvar, 
                           dimnames = list("Threshold",nameList), free=TRUE),
                  mxExpectationNormal(covariance="R", means="M", dimnames=nameList, 
                                      thresholds="ThresholdsClass1"),
                  mxFitFunctionML(vector=TRUE))

class2 <- mxModel("Class2", 
                  mxMatrix("Iden", name = "R", nrow = nvar, ncol = nvar, free=FALSE),
                  mxMatrix("Full", name = "M", nrow = 1, ncol = nvar, free=FALSE),
                  mxMatrix("Full", name = "ThresholdsClass2", nrow = 1, ncol = nvar, 
                           dimnames = list("Threshold",nameList), free=TRUE),
                  mxExpectationNormal(covariance="R", means="M", dimnames=nameList,
                                      thresholds="ThresholdsClass2"),
                  mxFitFunctionML(vector=TRUE))

# Define the model

lcamodel <- mxModel(model = "lcamodel", class1, class2, mxData(vars, type="raw"))

# Create class membership probabilities, constrain them to be in the range 0-1, and make their sum equal 1.0
mxMatrix("Full", name = "ClassMembershipProbabilities", nrow = nclass, ncol = 1, free=TRUE, labels = c(paste("pclass", 1:nclass, sep=""))) 
mxBounds(c(paste("pclass", 1:nclass, sep="")),0,1) 
mxMatrix("Iden", nrow = 1, name = "constraintLHS") 
mxAlgebra(sum(ClassMembershipProbabilities), name = "constraintRHS") 
mxConstraint(constraintLHS == constraintRHS)

# Define the objective function
mxAlgebra(-2*sum(freq * log(pclass1%x%Class1.objective + pclass2%x%Class2.objective)), name="lca")

mxFitFunctionAlgebra("lca")

# Run the job
model <- mxRun(lcamodel, suppressWarnings=TRUE)
summary(model)
model$matrices

# Check results against those hard-coded from old Mx:
negativeThresholds <- as.vector(mxEval(Class1.ThresholdsClass1, model))
positiveThresholds <- as.vector(mxEval(Class2.ThresholdsClass2, model))
if (negativeThresholds[[1]] > 0) {
  temp <- negativeThresholds
  negativeThresholds <- positiveThresholds
  positiveThresholds <- temp
}

```

Part 3: LCA using tidySEM

```{r}
#adjust this code
# The argument `classes` indicates which class solutions should be estimated (e.g., 1 through 6).
# The argument `variances` specifies whether variances should be `"equal"` or `"varying"` across classes.
# The argument `covariances` specifies whether covariances should be constrained to `"zero"`, `"equal"` or `"varying"` across classes.
tidySEM::mx_profiles(classes, variances, covariances)
```

After all models have been estimated, the function `tidySEM::table_fit()` can be used to obtain a model fit table suitable for determining the optimal model according to best practices.
Note however that this table does not include the bootstrapped likelihood ratio test (BLRT) by default,
because this test is very computationally expensive.
It is recommended to use the function `tidySEM::BLRT()` to compare a shortlist of likely candidate models based on other fit indices.

>>>>

---
title: "LCA in R using tidySEM"
output: html_notebook
---
This is an example of an exploratory LCA using `tidySEM`.

# Loading the Dataset

In this example, we use *data_mix_ordinal*, 
a simulated data for mixture modelling with ordinal indicators. 
This dataset is built-in to `tidySEM`.
For more information about the dataset, 
type `?tidySEM::data_mix_ordinal` into the R console
after loading the `tidySEM` using library().

We load the dataset and convert the indicators into ordered factors.

```{r, include = TRUE}
library(tidySEM) # loading tidySEM
df <- data_mix_ordinal  # dataset
df[1:4] <- lapply(df, ordered) # indicators as ordered factors
```

# Exploring the Data

An important step to preceding any statistical analysis is data exploration.
Here we use `tidySEM::descriptives()` 
to describe the dataset we are using.

```{r}
tidySEM::descriptives(df) # descriptives
```
As we can see, the output includes various descriptives of our dataset.
Special attention should be paid to examining the pattern of missingness,
as discussed in the *Handling missing data* section of this paper.
In our example, we see that there are no missing values,
hence we proceed with our analyses.

# Conducting Latent Class Analysis

Before we fit a series of LCA models, we set a random seed
using set.seed().
The seed number can be any digit.
This is an important step as 
there is some inherent randomness in the LCA computations,
and having the same seed number ensures that
two separate researcher obtain exactly the same results
when fitting LCA models.

Finally, we reach the step of fitting LCA models.
To do so, we use `tidySEM::mx_lca()`
which takes data and number of classes as inputs.
In our example, we want to fit 1 to 4 class solutions
and compare their output.
Depending on your computer's computational power,
this might take a while.

```{r}
set.seed(123) # setting seed 
res <- mx_lca(data = df, classes = 1:4) # fitting LCA 
```
# Class enumeration

In class enumeration, 
we want to compare a sequence of LCA models fitted to the data.
To do so, we use `tidySEM::table_fit()` 
with the results object as the input.
This function creates the model fit table formatted for publication.
However, it contains a lot of information on each of the four models.
It may be more helpful to focus on a subset of model fit indices
and classification diagnostics which we do in the subsequent step.

```{r}
fit_table <- table_fit(res) # model fit table
fit_table[ , c("Name", "LL", "AIC", "BIC", "Entropy", "prob_min", "prob_max", "n_min", "n_max")] # a selection from the model fit table
```
The model fit table is important for deciding on the final class solution.
Our subset of fit indices and classification diagnostics includes:

**Name** indicates the $K$-class solution.
In our case, we chose to fit 1 - 4 class solutions,
each is represented by its row in the table.
**LL** is the -2*log-likelihood of each model.
**AIC** and **BIC** are the Akaike 
and the Bayesian Information Criterion, respectively.
**prob_min** and **prob_max** are 
the lowest and the highest posterior class probability
by most likely class membership.
**n_min** and **n_max** are the lowest and the highest
class proportion based on the posterior class probabilities.

In this paper, we discussed several possible strategies 
to select the final class solution.

Here, we apply our own. 
First, we examine the -2*log likelihood 
which falls successively with each added class.

Then, we look at the BIC value
which is fairly close for the one, two, and three-class solutions,
but high for the four-class solution.

As previously stated, classification diagnostics should not be used for model selection,
but they can be used to disqualify certain solutions because they are uninterpretable.
We see that prob_min for the four-class solution is low,
therefore we disqualify this solution.

Out of the remaining three solutions,
we notice that entropy is the highest for the three-class solution,
and it has a satisfactory prob_min and n_min.
Based on this, we can retain the three class solution in model selection.
Note that entropy for the one-class solution will always equal to one,
as it is 100% true that every case is in that class.
Based on the low entropy of the two-class solution,
we eliminate this model.

Finally, when comparing the one and three-class solutions,
we consult the BIC which tells us that the added complexity
of having three classes still explains the data better
than a one-class solution.
Therefore, we select the three-class solution as our final-class solution.

# Interpreting Final Class Solution

To aid our understanding of the final class solution,
we use `ggplot2::plot_prob()` 
with the results of the three-class model as the input.
The resulting graph shows response patterns 
on all the indicators for each group.

If we want to know the probability of each response option's endoresement 
for each class, we can use `tidySEM::table_prob`.
These are thresholds for ordinal dependent variables in the probability scale.

```{r}
library(ggplot2)
plot_prob(res[[3]]) # plotting the response patterns for the final class solution
table_prob(res[[3]]) # tabulating the response patterns for the final class solution
```
In the plot, we can see the distributions each of the response probabilities on the indicators
for each of the three classes.
For instance, we see in Class 1 the most common response to u2 is 2,
while in Class 2 and Class 3 this is 0.
We can also see that response 1 is a rare response not forming the majority in 
any class.
Class 2 distinguishes itself because the majority scores the response 0 category of u3 and u4,
while in Class 1 and 2 this is not the case.
Class 3 distinguishes itself because the most common response to u3 and u4 is 3.

We can also interpret the response patterns numerically.

@@

Another step is to extract posterior class probabilities.
This is done by the use of `TidySEM::class_prob`
with the results of the final class solution as the input.

```{r}
probs <- class_prob(res[[3]]) # extracting the posterior class probabilities of the final class solution
probs$mostlikely.class # posterior class probabilities by most likely class membership
probs$individual # individual posterior class probabilities
```

BACKUP TUTORIAL (OLD)

This used to be in the tutorial section of the paper.
I store it here for backup, but I don't think it is longer relevant.

<!-- 5. Tutorial -->
<!-- 5.1 ‘mclust’ -->
<!-- To run LPA using the ‘mclust’ package, the Mclust() function can be used. This function -->
<!-- requires data as its minimum input. In addition, the number of clusters to fit (G) and model -->
<!-- variants to fit and select from (modelNames) can be entered (defaults are: G=1-9 and all -->
<!-- fourteen model variants). In many cases, we may want to consider a more limited range of -->
<!-- model variants to with different numbers of classes. For multivariate data, a total of fourteen -->
<!-- options are available (see help("mclustModelNames") for an overview). Each model variant -->
<!-- is estimated for each value of the specified range of G using Expectation Maximization (EM), -->
<!-- using the Bayesian Information Criterion (BIC) to compare different model variants and -->
<!-- select the optimal one. -->
<!-- Conducting a latent variable model or mixture analysis with Mplus or comparable software -->
<!-- usually entails fitting models with different numbers of classes (that are otherwise configured -->
<!-- similarly) and to compare the fit between these models to select the best model. The -->
<!-- default approach taken in the ‘mclust’ package is slightly different in that the main modeling -->
<!-- command Mclust() fits multiple models for all possible combinations of the specified number -->
<!-- of classes and model configurations . The eventual output of ‘mclust’ is the BIC-selected best -->
<!-- model variant with the optimal combination of number of classes and model configuration. -->
<!-- An advantage of the ‘mclust’ approach is that it is efficient and flexible: given a number of -->
<!-- classes, the best model configuration is selected straight away and we do not have to run all -->
<!-- model combinations one by one. This makes the approach especially suitable for exploratory -->
<!-- analyses. However, like stated above, this approach is somewhat less usual in some fields, -->
<!-- where estimation of LPA is approached in a more confirmatory fashion. In addition, it does -->
<!-- take some control away from the researcher, who may want to primarily focus on selecting -->
<!-- the number of classes, given a particular , theory-based model configuration that is kept -->
<!-- constant throughout analyses, which is often the chosen strategy in software like Mplus and -->
<!-- made easily available through the ‘tidyLPA’ package. -->
<!-- Here, both approaches are shown: (1) the ‘mclust’ aproach and (2) the ‘tidyLPA’ approach. -->

<!--                                               5 -->
<!-- 5.1.1 The ‘mclust’ approach -->
<!-- In this approach, we run consecutive models with increasing numbers of classes (G=1 to -->
<!-- G=9), for each model we let the package fit the four above described model variants (“EEI”, -->
<!-- “EEE”, “VVI” and “VVV”) and select the best-fitting one. We start by loading the package -->
<!-- (library(mclust)) and creating an object mnames that contains the names of the four models -->
<!-- we want to be fitted. Next, we use the Mclust() function to fit the models: -->
<!-- library(mclust) -->

<!-- ## Package 'mclust' version 5.4.7 -->
<!-- ## Type 'citation("mclust")' for citing this R package in publications. -->
<!-- mnames <- c("EEI", "EEE", "VVI", "VVV") -->

<!-- # Fit 1-5 class model -->
<!-- mod_g1_9 <- Mclust(dat1[, 1:10], modelNames = mnames) -->

<!-- We can first look at the optimal number of classes and the optimal model variant that were -->
<!-- selected, using the following code: -->
<!-- # Opimal number of classes -->
<!-- mod_g1_9$G -->

<!-- ## [1] 3 -->
<!-- # Optimal model variant -->
<!-- mod_g1_9$modelName -->

<!-- ## [1] "EEI" -->
<!-- This shows us that a 3-class EEI model was selected as the best model based on the BIC. We -->
<!-- can get a better perspective of this model’s performance if we compare it to the other fitted -->
<!-- models. We can do this by taking a closer look at the other models’ BIC values: -->
<!-- mod_g1_9$BIC -->

<!-- ##   Bayesian Information Criterion (BIC): -->
<!-- ##           EEI       EEE       VVI       VVV -->
<!-- ##   1 -14661.61 -12986.57 -14661.61 -12986.57 -->
<!-- ##   2 -13275.65 -12611.60 -13095.82 -12678.12 -->
<!-- ##   3 -12142.11 -12290.68 -12163.08 -12731.18 -->
<!-- ##   4 -12178.44 -12330.04 -12241.21 -12970.57 -->
<!-- ##   5 -12223.14 -12379.96 -12320.47 -13231.35 -->
<!-- ##   6 -12264.04 -12423.72 -12406.59 -13522.40 -->
<!-- ##   7 -12269.39 -12451.25 -12497.03 -13748.91 -->
<!-- ##   8 -12298.30 -12483.35 -12568.62 -13980.82 -->
<!-- ##   9 -12347.11 -12529.40 -12645.50 -14269.34 -->
<!-- ## -->

<!--                                              6 -->
<!-- ## Top 3 models based on the BIC criterion: -->
<!-- ##     EEI,3     VVI,3     EEI,4 -->
<!-- ## -12142.11 -12163.08 -12178.44 -->
<!-- mod_g1_9$loglik -->

<!-- ## [1] -5951.275 -->
<!-- mod_g1_9$df -->

<!-- ## [1] 42 -->
<!-- Here, we can see the BICs for all fitted models (in this case 1-9 classes and 4 model variants: -->
<!-- 36 models in total). We can see that the closest contenders were a 3-class model with a VVI -->
<!-- configuration (with variances allowed to vary both within and between classes) and a 4-class -->
<!-- EEI model. -->
<!-- Note that the BIC values are negative, and that values that are more negative are considered -->
<!-- to fit poorer than values closer to 0. This may strike some users as odd, given that in -->
<!-- many modeling applications, we are used to consider lower BIC values to indicate better -->
<!-- fit. However, the ‘mclust’ approach of maximizing the BIC is correct for the used modeling -->
<!-- approach (see e.g., Fraley & Raftery, 2003; Banfield & Raftery, 1993). For theoretical reasons, -->
<!-- the BIC is calculated in’mclust’ as: -->

<!--                            BIC = 2(Loglikelihood) - df (log(n)) -->

<!-- where df is the number of parameters (degrees of freedom) and n is the sample size. We can -->
<!-- see, that here, higher log likelihood values lead to a higher BIC. For the best fitting model in -->
<!-- our example, the log likelihood is -5951.275, the number of parameters is 42 and the sample -->
<!-- size is 300. If we plug these into the formula, we get the following BIC: -->

<!--                      BIC = 2(-5951.275) - 42(log(300)) = -12142.11 -->

<!-- As expected, the obtained value indeed corresponds to the BIC we got in the model output -->
<!-- (see above). Other software packages, such as Mplus and ‘tidyLPA’ (see below) use a slightly -->
<!-- different formula to calculate the BIC, where higher log likelihood values lead to lower BIC -->
<!-- values: -->

<!--                            BIC = df (log(n)) - 2(Loglikelihood) -->

<!-- When we plug the best-model values into this formula, we get: -->

<!--                       BIC = 42(log(300)) - 2(-5951.275) = 12142.11 -->

<!-- We can see that the absolute BIC value is the same, irrespective of the used formula, with -->
<!-- only the sign differing between the two approaches. -->

<!--                                               7 -->
<!-- Now, let us continue by taking a closer look at the class sizes and the mean values and -->
<!-- variances of the ten variables for the three identified classes. Because the best model has and -->
<!-- EEI configuration, printing out a single variance matrix from one class will suffice, because -->
<!-- the variances are the same across classes (this is different in cases where the optimal model -->
<!-- allows for between-class differences in (co)variances: -->
<!-- # tabulate class-membership numbers -->
<!-- table(summary(mod_g1_9)$classification) -->

<!-- ## -->
<!-- ##   1   2      3 -->
<!-- ## 100 110     90 -->
<!-- # display the means per class -->
<!-- mod_g1_9$parameters$mean -->

<!-- ##             [,1]         [,2]        [,3] -->
<!-- ##   var1 2.023572      6.083019   2.0304980 -->
<!-- ##   var2 3.056837      6.937785   4.8997993 -->
<!-- ##   var3 1.718166      1.107873   7.9385872 -->
<!-- ##   var4 6.086727      2.854402   6.9657536 -->
<!-- ##   var5 7.101109      2.225675   4.7416266 -->
<!-- ##   var6 3.691936      8.081570   2.2469977 -->
<!-- ##   var7 5.009819      2.817889   0.8862725 -->
<!-- ##   var8 8.316894      8.080891   3.9350536 -->
<!-- ##   var9 1.978136      1.896294   9.1537131 -->
<!-- ##   var10 1.191249     5.142590   8.0332628 -->
<!-- # select the first of 3 10-by-10 variance matrices -->
<!-- mod_g1_9$parameters$variance$sigma[1:10, 1:10, 1] -->

<!-- ##               var1       var2       var3       var4       var5       var6       var7       var8 -->
<!-- ##   var1    2.539537   0.000000   0.000000   0.000000   0.000000   0.000000   0.000000   0.000000 -->
<!-- ##   var2    0.000000   2.117941   0.000000   0.000000   0.000000   0.000000   0.000000   0.000000 -->
<!-- ##   var3    0.000000   0.000000   2.509035   0.000000   0.000000   0.000000   0.000000   0.000000 -->
<!-- ##   var4    0.000000   0.000000   0.000000   2.812796   0.000000   0.000000   0.000000   0.000000 -->
<!-- ##   var5    0.000000   0.000000   0.000000   0.000000   2.647399   0.000000   0.000000   0.000000 -->
<!-- ##   var6    0.000000   0.000000   0.000000   0.000000   0.000000   2.694156   0.000000   0.000000 -->
<!-- ##   var7    0.000000   0.000000   0.000000   0.000000   0.000000   0.000000   2.368379   0.000000 -->
<!-- ##   var8    0.000000   0.000000   0.000000   0.000000   0.000000   0.000000   0.000000   2.540704 -->
<!-- ##   var9    0.000000   0.000000   0.000000   0.000000   0.000000   0.000000   0.000000   0.000000 -->
<!-- ##   var10   0.000000   0.000000   0.000000   0.000000   0.000000   0.000000   0.000000   0.000000 -->
<!-- ##               var9      var10 -->
<!-- ##   var1    0.000000   0.000000 -->
<!-- ##   var2    0.000000   0.000000 -->
<!-- ##   var3    0.000000   0.000000 -->
<!-- ##   var4    0.000000   0.000000 -->

<!--                                                8 -->
<!-- ##   var5    0.000000   0.000000 -->
<!-- ##   var6    0.000000   0.000000 -->
<!-- ##   var7    0.000000   0.000000 -->
<!-- ##   var8    0.000000   0.000000 -->
<!-- ##   var9    2.509146   0.000000 -->
<!-- ##   var10   0.000000   2.204457 -->
<!-- From this output, we can see that the three classes have class-sizes of 100, 110 and 90, -->
<!-- respectively. These sizes along with the patterns of class-specific mean variable scores -->
<!-- correspond with those that served as the input for the simulation. As specified, the variances -->
<!-- vary slightly across variables. -->
<!-- The class allocation in LPA is probabilistic in nature: each subject in the data is assigned a -->
<!-- probability for each of the estimated classes, based on their pattern of scores on the input -->
<!-- variables. These probabilities can be inspected in the z-matrix (here: mod_g1_9$z). Subjects -->
<!-- can be allocated to one of the classes based on their highest class-probability. These posterior -->
<!-- allocations can be found in the classification matrix (here: mod_g1_9$classification). -->
<!-- These class-allocations were tabulated above to evaluate class sizes. -->
<!-- Note that class-allocation in this way only yields useful classifications if the patterns of -->
<!-- class-probabilities allow for allocation of each subject to a single class with sufficient certainty. -->
<!-- In case of too much uncertainty, using the classification is not advised. We can inspect the -->
<!-- uncertainty of allocation for all subjects (here: mod_g1_9$uncertainty). This gives us a list -->
<!-- of uncertainties for all subjects. Next, we can evaluate the extent of uncertainty by looking -->
<!-- at the maximum uncertainty or, for instance, the averaged uncertainty across subjects: -->
<!-- max(mod_g1_9$uncertainty) -->

<!-- ## [1] 0.003852604 -->
<!-- mean(mod_g1_9$uncertainty) -->

<!-- ## [1] 2.607299e-05 -->
<!-- Here, the uncertainty is atypically low because of the used idealized data. In many cases, -->
<!-- uncertainty is higher and it may be of interest to investigate it further, for instance, by -->
<!-- looking at (e.g., plotting or tabulating) the uncertainty per class: -->
<!-- cprob <- cbind(mod_g1_9$z, mod_g1_9$classification) -->
<!-- cprob <- as.data.frame(cprob) -->
<!-- colnames(cprob) <- c("prob (class 1)", "prob (class 2)", "prob (class 3)", "class") -->
<!-- aggregate(cprob[, 1:3], list(cprob$class), mean) -->

<!-- ##   Group.1 prob (class 1) prob (class 2) prob (class 3) -->
<!-- ## 1       1   9.999824e-01   1.761611e-05   3.350070e-10 -->
<!-- ## 2       2   5.486300e-05   9.999451e-01   7.192849e-08 -->
<!-- ## 3       3   2.591358e-08   1.675400e-07   9.999998e-01 -->
<!-- Here we can see that for each class, the mean probability of membership of that class was -->

<!--                                                  9 -->
<!-- very high and probabilities for the other classes were very low. Again, these values reflect the -->
<!-- idealized data; in many cases, more differentiated class-probability patterns may be observed. -->

<!-- 5.1.2 ‘tidyLPA’ -->
<!-- When running ‘mclust’ models from the ‘tidyLPA’ package, the estimations are approached -->
<!-- a little differently. By default, the number of models that can be estimated is restricted to -->
<!-- six relatively common variants (see help(tidyLPA::estimate_profiles) for details). Of -->
<!-- these models, four can be estimated with ‘mclust’. These models differ in terms of how the -->
<!-- variances and covariances are allowed to vary or constrained to be equal across classes and -->
<!-- whether covariances are or are not fixed to zero within classes. These models have each been -->
<!-- allocated a number (1-6) in ‘tidyLPA’. See the table below how these numbers correspond to -->
<!-- the above mentioned model configurations. -->
<!--  Model variant    model number -->
<!--  EEI              1 -->
<!--  EEE              3 -->
<!--  VVI              2 -->
<!--  VVV              6 -->
<!-- Models four and five are different and can only be fit when R is interfacing with Mplus using -->
<!-- the ‘MplusAutomation’ package. This is outside the scope of this tutorial. -->
<!-- To fit an LPA model in ‘tidyLPA’, we use the estimate_profiles() function. Here, we need -->
<!-- to enter the data.frame to be used (df) and the number of classes to estimate (n_profiles). -->
<!-- To determine what kind of model configuration to estimate, the authors have provided two -->
<!-- different ways in the estimate_profiles() command. In the first approach, we simply tell what -->
<!-- model number (see Table 2) we want to estimate using the models= argument For instance, if -->
<!-- we want to estimate LPA models with 1 to 9 classes, with an EEI configuration, we can use: -->
<!-- <!-- suppressMessages(library(tidyLPA)) --> -->
<!-- <!-- suppressMessages(mod_1c_v1 <- estimate_profiles(df = dat1[1:10], n_profiles = 1:9, --> -->
<!-- <!--     models = 1)) --> -->

<!-- The package issues a standard message by default notifying us that we use the models -->
<!-- argument and that the variances and covariances arguments are ignored. Here, we suppress -->
<!-- the message. The output shows us the model-fit information for the 1 to 9-class EEI models: -->
<!-- <!-- mod_1c_v1 --> -->

<!-- ## tidyLPA analysis using mclust: -->
<!-- ## -->
<!-- ## Model Classes AIC       BIC      Entropy            prob_min   prob_max   n_min   n_max   BLRT_p -->
<!-- ## 1      1       14587.53 14661.61 1.00               1.00       1.00       1.00    1.00 -->
<!-- ## 1      2       13160.84 13275.65 1.00               1.00       1.00       0.30    0.70    0.01 -->
<!-- ## 1      3       11986.55 12142.11 1.00               1.00       1.00       0.30    0.37    0.01 -->
<!-- ## 1      4       11982.14 12178.44 0.91               0.79       1.00       0.14    0.33    0.05 -->
<!-- ## 1      5       11986.09 12223.14 0.89               0.75       1.00       0.06    0.33    0.30 -->

<!--                                               10 -->
<!-- ##   1      6         11986.25    12264.04   0.84      0.78       1.00        0.08   0.30    0.16 -->
<!-- ##   1      7         11950.86    12269.39   0.81      0.77       0.91        0.08   0.17    0.01 -->
<!-- ##   1      8         11939.04    12298.30   0.82      0.79       0.92        0.08   0.17    0.02 -->
<!-- ##   1      9         11947.10    12347.11   0.82      0.76       0.92        0.06   0.16    0.57 -->
<!-- Here, we can see that the BIC takes on different values compared to ‘mclust’, with lower -->
<!-- rather than higher values indicating best fit (see explanation above). Instead of drawing the -->
<!-- fit indices (e.g., BIC, AIC) directly from the ‘mclust’ package, ‘tidyLPA’ only draws the ‘raw’ -->
<!-- log likelihood values and posterior probabilities from the ‘mclust’ output and (re)calculates -->
<!-- the BIC, AIC, entropy etc. to mirror as well as possible those provided in Mplus. In addition, -->
<!-- p-values of the the bootstrapped likelihood ratio test (BLRT) are given by default. These -->
<!-- p-values indicate for each k-class model whether adding the kth class adds significantly to -->
<!-- model fit. For instance, the BLRT_p of 0.04 for the 4-class model indicates that adding -->
<!-- a fourth class led to an improvement of model fit compared to the 3-class model that was -->
<!-- just significant when using an alpha of 0.05. When using ’mclust’ as a stand-alone package, -->
<!-- the BLRT is not calculated by default, but can be obtained with the mclustBootstrapLRT -->
<!-- function. -->
<!-- Next, we can run similar commands, with different values for the models argument. It is -->
<!-- also possible to estimate more than one model variant in a single run by providing a vector -->
<!-- of model numbers (e.g., models=c(1,6)). -->
<!-- The second approach to determine the model variant to be estimated is by using the variances -->
<!-- and covariances arguments in the estimate_profiles command. The variances argument -->
<!-- can have the values "equal" (i.e. variable variances constrained to be equal across classes) or -->
<!-- "varying" (i.e. variances allowed to vary across classes). The covariances argument can -->
<!-- have the values: "zero" (i.e. all variable covariances fixed to zero), "equal" (i.e. covariances -->
<!-- constrained to be equal across classes) or "varying" (i.e. covariances allowed to vary across -->
<!-- classes). Now, if we again want to estimate LPA models with 1 to 9 classes, with an EEI -->
<!-- configuration, we can use: -->
<!-- <!-- mod_1c_v2 <- estimate_profiles(df = dat1[1:10], n_profiles = 1:9, variances = "equal", --> -->
<!-- <!--     covariances = "zero") --> -->
<!-- <!-- mod_1c_v2$model_1_class_2$fit --> -->

<!-- <!-- ##         Model           Classes        LogLik                   AIC              AWE --> -->
<!-- <!-- ## 1.000000e+00       2.000000e+00 -6.549418e+03          1.316084e+04     1.354347e+04 --> -->
<!-- <!-- ##           BIC              CAIC           CLC                   KIC            SABIC --> -->
<!-- <!-- ## 1.327565e+04       1.330665e+04 1.310084e+04           1.319484e+04     1.317734e+04 --> -->
<!-- <!-- ##           ICL           Entropy      prob_min              prob_max            n_min --> -->
<!-- <!-- ## -1.327568e+04      9.996750e-01 9.999484e-01           9.999990e-01     3.000000e-01 --> -->
<!-- <!-- ##         n_max          BLRT_val        BLRT_p --> -->
<!-- <!-- ## 7.000000e-01       1.448699e+03 9.900990e-03 --> -->
<!-- As expected, the output is exactly the same as the one we got when using the models -->
<!-- argument. Now, if we want to, we can continue fitting other model variants using either -->
<!-- approach. Next, we can manually compare the models in terms of their BIC and AIC values -->

<!--                                               11 -->
<!-- or we can use compare_solutions to do this in an automated fashion. Interestingly, the -->
<!-- best model can be selected based on the integrated information about several fit indices -->
<!-- (analytic hierarchy process; see help(tidyLPA::AHP)) for more details). We obtain the model -->
<!-- comparison with the following code. -->
<!-- <!-- comp <- suppressWarnings(compare_solutions(mod_1c_v1)) --> -->
<!-- <!-- comp$fits --> -->
